---
title: "Analyze multimodal clustering from Specter"
author: "Canzar Lab"
date: "4/14/2020"
#output: html_document
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load library and data

```{r message=FALSE}
library(Seurat)
library(ggplot2)
library(mclust)
library(plyr)
library(dplyr)    
```
Next we load two count matrices : one for the RNA measurements, and one for the antibody-derived tags (ADT).

```{r}
selected_cells <- read.csv(file = "data/pbmc_nodoublet_minPts50_selectedcells.csv", sep = ",", 
                           header = TRUE, row.names = 1)

pbmc.rna <- as.sparse(read.csv(file = "data/pbmc_rna.csv", sep = ",", 
                               header = TRUE, row.names = 1))
pbmc.rna <- pbmc.rna[, rownames(selected_cells)] # remove doublet

# Load in the ADT UMI matrix
adt <- read.csv(file = "data/pbmc_adt.csv", sep = ",", 
                header = TRUE, row.names = 1)
pbmc.adt <- as.sparse(adt[1:49, ])
pbmc.adt <- pbmc.adt[ ,rownames(selected_cells)]
```
## Create Seurat object and preprocess the data
### Preprocessing mRNA
```{r warning=FALSE}
pbmc <- CreateSeuratObject(counts = pbmc.rna)
# standard log-normalization
pbmc <- NormalizeData(pbmc)
# choose variable features
pbmc <- FindVariableFeatures(pbmc)
# standard scaling
pbmc <- ScaleData(pbmc)
# Run PCA, select 20 PCs for tSNE visualization and graph-based clustering
pbmc <- RunPCA(pbmc, verbose = FALSE)
# Run t-SNE on mRNA
pbmc <- RunTSNE(pbmc, dims = 1:20, method = "FIt-SNE")
# DimPlot(pbmc, label = TRUE) + NoLegend()
```


### Preprocessing ADT

```{r warning=FALSE}
pbmc[["ADT"]] <- CreateAssayObject(counts = pbmc.adt)

# Use a centered log-ratio (CLR) normalization
pbmc <- NormalizeData(pbmc, assay = "ADT", normalization.method = "CLR")
pbmc <- ScaleData(pbmc, assay = "ADT")
# Set ADT data as default. 
DefaultAssay(pbmc) <- "ADT"
pbmc <- RunPCA(pbmc, features = rownames(pbmc), reduction.name = "pca_adt", reduction.key = "pca_adt_", 
               verbose = FALSE)
#DimPlot(pbmc, reduction = "pca_adt")
# Compute a standard euclidean distance matrix, which is used for visualization and clustering ADT count
adt.data <- GetAssayData(pbmc, slot = "data")
adt.dist <- dist(t(adt.data))

# Compute tSNE using our distance matrix defined only on ADT (protein) levels.
pbmc[["tsne_adt"]] <- RunTSNE(adt.dist, assay = "ADT", reduction.key = "adtTSNE_")

#DimPlot(pbmc, reduction = "tsne_adt") + NoLegend()
```

## Read and visualize clusters from Specter

```{r}
sce_labels <- read.csv("output/pbmc_specter_clustering.csv", header = F)
sce_labels <- as.factor(sce_labels$V1)
names(sce_labels) <- pbmc.adt@Dimnames[[2]]

pbmc[["sceClusterID"]] <- sce_labels
DefaultAssay(pbmc) <- "RNA"
Idents(pbmc) <- sce_labels
```

```{r  fig26, fig.height = 6, fig.width = 12, warning=FALSE}
#Plot clusters on RNA and ADT t-SNE. 
tsne_rnaClusters <- DimPlot(pbmc, reduction = "tsne", group.by = "sceClusterID") + NoLegend()
tsne_rnaClusters <- tsne_rnaClusters + ggtitle("Specter's clusters on RNA") + theme(plot.title = element_text(hjust = 0.5))
tsne_rnaClusters <- LabelClusters(plot = tsne_rnaClusters, id = "sceClusterID", size = 4)

tsne_adtClusters <- DimPlot(pbmc, reduction = "tsne_adt", group.by = "sceClusterID", pt.size = 0.5) + NoLegend()
tsne_adtClusters <- tsne_adtClusters + ggtitle("Specter's clusters on ADT") + theme(plot.title = element_text(hjust = 0.5))
tsne_adtClusters <- LabelClusters(plot = tsne_adtClusters, id = "sceClusterID", size = 4)
CombinePlots(plots = list(tsne_rnaClusters, tsne_adtClusters), ncol = 2)
```

## Running DE test to assign cell type identity to clusters
### DE test on RNA
__Note that gene name starts with a prefix "hg19".__ 
```{r message=FALSE}
# We need to switch to RNA
DefaultAssay(pbmc) <- "RNA"
specter.rna.markers <- FindAllMarkers(pbmc, max.cells.per.ident = 100, min.diff.pct = 0.3, only.pos = TRUE)
# Extract top 10 marker genes for each cluster
rna.top10 <- specter.rna.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
# head(rna.top10, n = 15)
```

You can plot heatmap on top markers genes

```{r message=FALSE}
# Plot heatmap on top marker ADTs
# DoHeatmap(pbmc, features = unique(rna.top10$gene), assay = "RNA", angle = 0) 
```


### DE test on ADT
```{r message=FALSE}
# Find protein markers for all clusters, and draw a heatmap
DefaultAssay(pbmc) <- "ADT"
adt.markers <- FindAllMarkers(pbmc, assay = "ADT", only.pos = TRUE)
# Extract ADT top 10 markers
adt.top10 <- adt.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
#head(adt.top10, n = 15)
```

Plot heatmap on 10 markers ADT, but it is not very informative. 

```{r  fig2aaa, fig.height = 10, fig.width = 14, warning=FALSE, message=FALSE}
# Plot heatmap on top marker ADTs
#DoHeatmap(pbmc, features = unique(adt.top10$gene), assay = "ADT", angle = 0) 
```

For PBMCs cells we have canonical markers to match the unbiased clustering to known cell types:

Cell Type     | Marker genes
------------- | -------------
Naive CD4+ T  | IL7R, CCR7, SELL	
Memory CD4+   | IL7R, S100A4, CD27
CD14+ Mono    | CD14, LYZ	
B             | MS4A1
CD8+ T        | CD8A
FCGR3A+ Mono  | FCGR3A, MS4A7	
NK            | GNLY, NKG7	
DC            | FCER1A, CST3	


### Look at the top marker genes and ADT, we get the following cluster IDs.

```{r}
new.cluster.ids <- c("CD4+CD27-DR+ T","CD8+CD27- T","CD8+CD27- T","CD8+CD27- T","CD4+CD27+ T", "CD4+CD27-DR- T","CD14+ Mono","FCGR3A+ Mono", "CD14+ Mono","CD8+CD27+ T","Naive CD4+ T","MK","NK","CD8+CD27- T","Effector CD8+ T","B") 
```

### Visualize on cluster IDs

```{r}
sce_labels <- mapvalues(sce_labels, from = levels(sce_labels), to = new.cluster.ids)
pbmc[["sceClusterID"]] <- sce_labels
Idents(pbmc) <- sce_labels
```
```{r  fig2, fig.height = 6, fig.width = 12, warning=FALSE}
#Plot clusters on RNA and ADT t-SNE. 
tsne_rnaClusters <- DimPlot(pbmc, reduction = "tsne", group.by = "sceClusterID") + NoLegend()
tsne_rnaClusters <- tsne_rnaClusters + ggtitle("Specter's clusters on RNA") + theme(plot.title = element_text(hjust = 0.5))
tsne_rnaClusters <- LabelClusters(plot = tsne_rnaClusters, id = "sceClusterID", size = 4)

tsne_adtClusters <- DimPlot(pbmc, reduction = "tsne_adt", group.by = "sceClusterID", pt.size = 0.5) + NoLegend()
tsne_adtClusters <- tsne_adtClusters + ggtitle("Specter's clusters on ADT") + theme(plot.title = element_text(hjust = 0.5))
tsne_adtClusters <- LabelClusters(plot = tsne_adtClusters, id = "sceClusterID", size = 4)
CombinePlots(plots = list(tsne_rnaClusters, tsne_adtClusters), ncol = 2)
```


## Cluster validation
### Similar celltypes as Specter

Confirm CD27+ vs CD27- subtypes (based on CD27)

```{r  fig1b, fig.height = 6, fig.width = 12, message=FALSE}
RidgePlot(pbmc, assay = "ADT", features = c("CD27"), ncol = 2)
```

Confirm DR+ DR- subtype based on genes "NKG7", "GZMA" (see Specter)

```{r  fig1bc, fig.height = 6, fig.width = 12, message=FALSE}
all_markers <- c("NKG7", "GZMA")
RidgePlot(pbmc, assay = "RNA", features = paste("hg19-", all_markers, sep=""), ncol = 2)
```














